## Motivation

I just learned about the boostrap, which appears to be a versatile and powerful technique. The purpose of this short notebook is to convince myself that it actually works. 

## Toy Example

Perhaps the easiest way to use the boostrap is to estimate the mean and variance of the sample mean. First let's simulate some data.

```{r}
n = 100 #sample size
mu = 5 #mean
sigma = 3 #sd

X = rnorm(n, mu, sigma) 
```

So here we have a random sample of `r n` points drawn from $\mathcal{N}(`r mu`,`r sigma^2`)$. Next we can take bootstrap samples from $\mathbf{X}$ to estimate the sample mean and its variance.

We expect (asymptotically) $\bar{X} \sim \mathcal{N}(`r mu`, `r (sigma^2)/n`)$. Let's see how well our bootstrap estimates hold up to the theory.

```{r}
B = 10000 #number of bootstrap samples
b_i = numeric(n) #to store each bootstrap sample
xbar_i = numeric(B) #to store the sample means


for (i in 1:length(xbar_i)) {
  b_i = sample(X, size = n, replace = TRUE) #sample n points from X with replacement
  xbar_i[i] = mean(b_i) # take sample mean of bootstrap sample
}
```

First we can estimate the mean of the sample mean
```{r}
mu_xbar = mean(xbar_i)
mu_xbar
```

Next we can estimate the variance of the sample mean
```{r}
sigma_xbar = var(xbar_i)
sigma_xbar
```

We see that the bootstrap estimated mean, `r mu_xbar`, and variance, `r sigma_xbar`, are remarkably close to the true theoretical asymptotic mean, `r mu`, and variance, `r (sigma^2)/n`.


We can also use the 2.5% and 97.5% percentiles to estimate a 95% confidence interval. 

```{r}
quantile(xbar_i, c(.025, .975))
```

Let's compare this to the confidence interval we would get from using the estimated variance.

```{r}
cat(paste0("(",round(mu_xbar-1.96*sqrt(sigma_xbar),6),", ",round(mu_xbar+1.96*sqrt(sigma_xbar),6),")"))
```

These are pretty close. 

I think I've proved to myself that despite its simplicity, the bootstrap does in fact work to characterize estimators very well. This example was, of course, trivial because we know the sampling distribution of the sample mean due to the Central Limit Theorem, but if the sampling distribution is unknown, the bootstrap is a great tool. 

